---
title: "Peer Commentary"
author: "Jimmy Erkens"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: cosmo
---
# Jimmy commentary

## Preliminaries

Solid! Loading numerous libraries in one function is a solid application from the initial documentation. This is a personal preference, but I generally prefer to use numerous `library()` arguments so that if there are any hiccups with loading packages, I can pinpoint the line compiling failed. But, this is just how my brain works I think.

## Challenge 1

Good job finding `str_split_1()`! This greatly simplifies our procedure and eliminates the necessity of an `unlist()` function. I will say in actual practice that variable names like `v1` or `v2` can be a bit meaningless and unclear. There's a lot of boring theory behind variable naming conventions but I generally keep it related to the variable being described. 

## Challenge 2 and 3

Nice comments! Makes it clear what's going on and as far as I can tell your takes on what's going on is correct.

## Challenge 4

Good job nesting these lists! Just to make the knit a bit cleaner, I'd personally only print the final `Primates` list but again personal preference.

## Challenge 5

Right on! It's clear what's going on with the code, and your comments are excellent. I also see you used the semicolon method to print out a variable (I honestly earned this from a comp stat prof and it expanded my mind).

# **Challenges faced**

1. For Challenge 1, I had a hard time figuring out how to extract every third element from the vector.
    + See comments and solution underneath
2. At first, I had a hard time interpreting the output of the matrix and figuring out where the outputted elements of the matrix were coming from
    + Re-reading module, referencing Ch3 in Book of R, comparing original matrix to extracted elements all helped me figure this out
3. For Challenge 4, I initially forgot to name the lists after making them so the output didn't look like the example for Strepsirhini
    + Looked fine after I added that in

# **Preliminaries**
Set working directory and load library
```{r preliminaries}
library(stringr, yaml)
```

# **Challenge 1**
First, create a vector of 1, word character strings comprising the first line of the novel Ulysses, by James Joyce: “Stately, plump Buck Mulligan came from the stairhead, bearing a bowl of lather on which a mirror and a razor lay crossed.” Then, extract every third element of this vector.

## First attempt
```{r challenge 1 attempt 1}
v1 <- "Stately, plump Buck Mulligan came from the stairhead, bearing a bowl of lather on which a mirror and a razor lay crossed."; v1 
#this makes entire sentence one character 

v1<- gsub("[[:punct:]]","",v1); v1 
#remove punctuation

v1 <- str_split(v1, " "); v1
#creates list containing character vectors

v1.1 <- v1[seq(from = 1, to = length(v1), by = 3)]; v1.1
#this isn't doing anything, try again sam!
```

## Second attempt
Using str_split_1 function instead of str_split <br>
* `str_split` takes character vector and returns a list <br>
* `str_split_1` takes single string and returns a character vector <br>
* I think my first attempt didn't work because i turned the character vector into a list before trying to extract every third element
```{r challenge 1 attempt 2}
v2 <- "Stately, plump Buck Mulligan came from the stairhead, bearing a bowl of lather on which a mirror and a razor lay crossed."; v2
# this creates a single string from the sentence

v2 <- gsub("[[:punct:]]","",v2); v2

v2 <- str_split_1(v2, " "); v2
#this splits the string up into character vectors

v2[seq(from = 1, to = length(v2), by = 3)]; v2
# this extracts every third character vector
```

# **Challenge 2**
```{r challenge 2}
m <- matrix(data = 1:80, nrow =8, ncol = 10, byrow = FALSE); m
x <- m[4,5]; x #extract element in row 4, column 5 and assign to new variable x

#extract 2nd, 3rd, 6th COLUMNS and assign to new variable x
x <- m[, c(2,3,6)]; x

#extract 6th to 8th ROW and assign to new variable x
x <- m[6:8, ]; x

#extract elements from [row 2, column 2] to [row 6, column 9] and assign to new variable x
x <- m[2:6, 2:9]; x
```

# **Challenge 3**
Construct a 4-dimensional, 400 element array (5 x 5 x 4 x 4) named a consisting of the numbers 400 to 1 (i.e., a descending series)
```{r challenge 3a}
a <- array(data = 400:1, dim = c(5,5,4,4)); a #outputs 16 matrices
```
Given this matrix, what would the following return:
```{r challenge3b}
a[1,1,1,2]
# returns value 300; comes from row 1 and column 1 (1,1) of matrix (1,2)

a[2,3,2,]
# returns [1] 364 264 164 64
# elements from row 2 and column 3 (2,3) of matrices (2, 1:4)

a[1:5, 1:5, 3, 3]
# returns elements from rows and columns 1:5, aka all elements of (3,3) matrix
```

# **Challenge 4**
Create a list representing this simplified primate taxonomy:
``` {r challenge 4}
# list for strepsirrhines, already provided
Lemuroidea <- c("Cheirogaleidae", "Lepilemuridae", "Indriidae", "Lemuridae", "Daubentoniidae")
Lorisoidea <- c("Lorisidae", "Galagidae")
Strepsirhini <- list(Lemuroidea, Lorisoidea) #nesting within streps
  names(Strepsirhini) <- list("Lemuroidea", "Lorisoidea") #naming the list
Strepsirhini #final output

# making lists for haplorhines
Platyrrhini <- c("Cebidae", "Atelidae", "Pitheciidae")
Catarrhini <- c("Cercopithecidae", "Hylobatidae", "Hominidae")
Tarsioidea <- c("Tarsiidae")

# nesting and namining lists for haplorhines
Anthropoidea <- list(Platyrrhini, Catarrhini)
  names(Anthropoidea) <- list("Platyrrhini", "Catarrhini")
Haplorhini <- list(Anthropoidea, Tarsioidea)
  names(Haplorhini) <- list("Anthropoidea", "Tarsioidea")
Haplorhini

# making final list for all primates
Primates <- list(Haplorhini, Strepsirhini)
  names(Primates) <- list("Haplorhini", "Strepsirhini")
Primates
```

# **Challenge 5**
```{r challenge 5a}
m3 <- matrix(data=c(3, 0, 1 ,23, 1, 2, 33, 1, 1, 42, 0, 1, 41, 0, 2), nrow=5, ncol=3, byrow=TRUE); m3
# storing these numbers into a 5x3 matrix that's filled in row-wise

m3_df <- data.frame(m3); m3_df 
# coerce matrix to data frame, can view data frame in output to see names of columns

m3_df$X2 <- as.logical(m3_df$X2) 
# coerce second column to be logical-valued

m3_df$X3 <- as.factor(m3_df$X3) 
# coerce third column to be factor-valued

str(m3_df) 
# this command shows data type for each variable in data frame!
```
